---
title: "ARMA-GARCH Estimation"
author: "Alexios Galanos"
date: "`r Sys.Date()`"
output: 
    rmarkdown::html_vignette:
        css: custom.css
        code_folding: show
vignette: >
  %\VignetteIndexEntry{ARMA GARCH Estimation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


*Since version 1.0.5 joint estimation of ARMA and GARCH is available.*

## The Model

The observed series $y_t$ is decomposed into a conditional mean $\mu_t$ and
an innovation $\varepsilon_t$:

$$
y_t = \mu_t + \varepsilon_t
$$

By default (`arma = c(0,0)`) $\mu_t$ reduces to either a constant $\mu$ or
zero, controlled independently by the `constant` argument to
`garch_modelspec`. Setting `arma = c(p,q)` instead activates a jointly
estimated ARMA($p,q$) mean equation, in which $\mu_t$ is time-varying and
follows the mean-adjusted recursion

$$
\mu_t = \mu + \sum_{i=1}^{p}\phi_i\left(y_{t-i}-\mu\right) + \sum_{j=1}^{q}\theta_j\varepsilon_{t-j}
$$

so that $\mu$ retains its interpretation as the unconditional mean of $y$,
and the AR feedback term is centered on $y_{t-i}-\mu$ rather than the raw
$y_{t-i}$. This is the same mean-adjusted form used in **rugarch**'s
ARFIMA-X mean equation.

Combining the two equations above, the innovation $\varepsilon_t$ is
obtained recursively as

$$
\varepsilon_t = y_t - \mu_t = \left(y_t-\mu\right) - \sum_{i=1}^{p}\phi_i\left(y_{t-i}-\mu\right) - \sum_{j=1}^{q}\theta_j\varepsilon_{t-j}
$$

which only requires $\varepsilon_{t-1},\ldots,\varepsilon_{t-q}$ from
previous time steps, so the whole recursion can be evaluated forward in a
single pass once $\mu$, $\phi_i$ and $\theta_j$ are known. Setting
`arma = c(0,0)` collapses both equations exactly to the constant-mean case,
$\varepsilon_t = y_t - \mu$, matching prior (pre 1.0.5) behavior with no
change in results.

It is $\varepsilon_t$ (rather than $y_t-\mu$) that is passed on to the
conditional variance equation of whichever GARCH flavor is chosen (see the
GARCH Models vignette for the flavor-specific variance equations), decomposed as

$$
\varepsilon_t = \sigma_t z_t, \qquad z_t \overset{\text{iid}}{\sim} D(0,1;\ldots)
$$

with $\sigma_t$ the conditional standard deviation implied by the chosen
GARCH variance equation, and $z_t$ an iid draw from the chosen conditional
distribution $D$ (normal, Student-t, skew distributions, etc.), with any
shape/skew parameters estimated jointly with everything else.

The combined pre-sample/burn-in length used to initialize this recursion is
$\max\left(\text{GARCH order}, \text{ARMA order}\right)$, so that e.g. an
ARMA(2,2) mean equation together with a GARCH(1,1) variance equation
initializes correctly; over that pre-sample the process is assumed to start
at its unconditional mean with zero shocks, i.e. $y_t = \mu$ and
$\varepsilon_t = 0$.

## Regressors in the Mean Equation (ARMAX)

Regressors can also enter the conditional mean via the `xreg` argument to
`garch_modelspec`, with coefficients named `tau1`, `tau2`, ... in the
`parmatrix` ($\tau_k$; the symbol $\xi$ is already used in this package
for the variance equation regressors). Two conventions are available
through `xreg_type`. The default, `"arma_errors"`, runs the ARMA
recursion on the regressor-adjusted deviations

$$
y_t - \mu - \sum_{k=1}^{s}\tau_k x_{k,t}
= \sum_{i=1}^{p}\phi_i\left(y_{t-i} - \mu - \sum_{k=1}^{s}\tau_k x_{k,t-i}\right)
+ \varepsilon_t + \sum_{j=1}^{q}\theta_j\varepsilon_{t-j}
$$

matching `stats::arima(xreg = )` and the sibling **tsarma** package, so
that $\tau_k$ is the long-run marginal effect of $x_{k,t}$. The
alternative, `"armax"` (the **rugarch** convention), adds the regressor
contribution to the conditional mean at time $t$ only,

$$
y_t = \mu + \sum_{i=1}^{p}\phi_i\left(y_{t-i}-\mu\right)
+ \sum_{k=1}^{s}\tau_k x_{k,t} + \varepsilon_t
+ \sum_{j=1}^{q}\theta_j\varepsilon_{t-j}
$$

so that $\tau_k$ is the impact effect and the long-run effect of a
sustained change is $\tau_k/\left(1-\sum_i\phi_i\right)$. The two
conventions differ only by the lagged-regressor term inside the AR
feedback and are algebraically identical whenever $p = 0$.

`xreg` must be an xts matrix aligned with `y`: same number of rows,
matching time index, no `NA`/`NaN`/`Inf` values, and full column rank
(including against the constant when `constant = TRUE`) - violations are
rejected at specification time. Because the recursion consumes a
regressor value at every step, `predict()` and `simulate()` require
future values (`newxreg` with `h` rows, `xreg` with `h + burn` rows
respectively), and `tsfilter()` requires `newxreg` covering the appended
observations; if the model was specified with `xreg` but no future
values are supplied, a zero matrix is substituted with a warning rather
than an error.

A short demo with a day-of-week effect in the Nikkei returns:

```{r}
library(tsgarch)
suppressMessages(library(xts))
data(nikkei)
nikkei <- xts(nikkei$value, as.Date(nikkei$index))
monday <- xts(as.numeric(weekdays(index(nikkei)) == "Monday"), index(nikkei))
colnames(monday) <- "monday"
spec_x <- garch_modelspec(nikkei, arma = c(1,1), xreg = monday,
                          xreg_type = "arma_errors", model = 'garch',
                          constant = TRUE, init = 'unconditional',
                          distribution = 'jsu')
mod_x <- estimate(spec_x)
coef(mod_x)["tau1"]
```

Forecasting requires the future regressor values:

```{r}
newx <- xts(matrix(as.numeric(weekdays(index(nikkei)[NROW(nikkei)] + 1:10) == "Monday"),
                   ncol = 1), index(nikkei)[NROW(nikkei)] + 1:10)
colnames(newx) <- "monday"
predict(mod_x, h = 10, newxreg = newx)$mean
```

### Which convention should I use?

Written out in levels for an AR(1) mean equation, `arma_errors` is

$$
y_t = \mu\left(1-\phi_1\right) + \phi_1 y_{t-1} + \tau_1 x_{1,t} - \phi_1\tau_1 x_{1,t-1} + \varepsilon_t
$$

and `armax` is the same expression without the final regressor lag. Both are
restrictions of the autoregressive distributed lag model
$y_t = c + \phi_1 y_{t-1} + \beta_0 x_{1,t} + \beta_1 x_{1,t-1} + \varepsilon_t$:
`arma_errors` imposes $\beta_1 = -\phi_1\beta_0$ and `armax` imposes
$\beta_1 = 0$. Neither is nested in the other.

The difference that matters in practice is propagation. Under `arma_errors`
a one period blip in a regressor shifts $y_t$ by $\tau_k$ and nothing
afterwards, because the ARMA dynamics apply only to the error; $\tau_k$ is
both the impact and the total effect. Under `armax` the same blip is carried
forward by the autoregressive term as $\tau_k, \phi_1\tau_k,
\phi_1^2\tau_k,\ldots$, so $\tau_k$ is the impact effect and
$\tau_k/\left(1-\sum_i\phi_i\right)$ the cumulative one. Event and calendar
dummies usually suit the former, covariates whose influence builds and
decays with the series the latter. With no AR terms the two are identical,
so the choice only matters when the mean equation has autoregressive
dynamics and the regressor is itself persistent.

Since `armax` with both $x_{k,t}$ and $x_{k,t-1}$ as separate `xreg` columns
spans the unrestricted model, the restriction each convention imposes can be
tested directly by fitting that larger model and inspecting $\hat\beta_1$.
See the GARCH Models vignette for the fuller discussion, including how
misspecifying the convention distorts the estimated persistence rather than
the regressor coefficient.

## Stationarity and Invertibility

For $\phi=\left(\phi_1,\ldots,\phi_p\right)$ and
$\theta=\left(\theta_1,\ldots,\theta_q\right)$ to be well defined ARMA
coefficients, the AR polynomial

$$
\Phi(z) = 1 - \phi_1 z - \cdots - \phi_p z^p
$$

must have all roots strictly outside the unit circle (stationarity), and
the MA polynomial

$$
\Theta(z) = 1 + \theta_1 z + \cdots + \theta_q z^q
$$

must have all roots strictly outside the unit circle (invertibility).
Equivalently, the *inverse* roots of both polynomials must lie strictly
inside the unit circle - this is exactly what the first panel of
`plot(object, type = "arma")` displays, and what `arma_inverse_roots()`
computes programmatically.

Rather than impose these two conditions as nonlinear inequality
constraints on $\phi_i$ and $\theta_j$ directly during optimization (which
would require either finite-difference Jacobians, or an autodiff-aware
eigendecomposition of a companion matrix), `tsgarch` instead reparameterizes
the raw quantities being optimized over as partial-autocorrelation-like
parameters $r^{ar}_i, r^{ma}_j \in \left(-1,1\right)$, and recovers $\phi$
and $\theta$ from them via the Durbin-Levinson / Jones recursion, which is
guaranteed by construction to always land in the stationarity/invertibility
region (Barndorff-Nielsen and Schou, 1973; Jones, 1980) - the same device
used internally by `stats::arima(..., transform.pars = TRUE)`.

Concretely, given raw parameters $r_1,\ldots,r_p\in\left(-1,1\right)$, the
AR coefficients are obtained by the forward recursion

$$
\phi^{(1)}_1 = r_1, \qquad
\phi^{(k)}_k = r_k, \qquad
\phi^{(k)}_i = \phi^{(k-1)}_i - r_k\,\phi^{(k-1)}_{k-i} \quad \left(i=1,\ldots,k-1\right)
$$

for $k=2,\ldots,p$, with the final AR coefficients given by
$\phi_i = \phi^{(p)}_i$. The same recursion applied to the *negated* raw
MA parameters, with the result negated again, yields MA coefficients
$\theta_j$ that satisfy invertibility, via the well-known AR/MA duality:

$$
\theta = -\,\text{DL}\left(-r^{ma}_1,\ldots,-r^{ma}_q\right)
$$

where $\text{DL}(\cdot)$ denotes the recursion above. Because this map is a
smooth (indeed, polynomial) function of the raw parameters, it differentiates
exactly under automatic differentiation, so **TMB** provides exact (not
finite-difference) Jacobians of the whole ARMA mean equation throughout
estimation, prediction, simulation and filtering. Only simple box
constraints $r^{ar}_i, r^{ma}_j \in \left(-1,1\right)$ (tightened slightly
to $\left(-0.995, 0.995\right)$ for numerical safety near the boundary) are
needed during optimization - no nonlinear stationarity/invertibility
constraint is ever evaluated or differentiated.

The raw parameters themselves (`arpacf`/`mapacf` in the model's
`parmatrix`) are not directly interpretable; the transformed $\phi$/$\theta$
coefficients are what `summary()` displays, and are also available via
`arma_coefficients()`.

## Diagnostic helpers and the "ARMA(1,1) trap"

Four helpers are exported for inspecting the estimated mean equation:

- `arma_coefficients()` returns the transformed AR and MA coefficients
  $\phi_i$ and $\theta_j$.
- `arma_inverse_roots()` returns the inverse roots of $\Phi(z)$ and
  $\Theta(z)$; all of them must lie inside the unit circle.
- `arma_irf()` computes the impulse-response function of the mean equation
  (how a one-unit shock to $\varepsilon_t$ propagates through $\mu_t$).
- `arma_near_cancellation()` flags pairs of AR and MA inverse roots that
  are closer than a chosen tolerance in the complex plane.

The first three are mostly self-explanatory, but the last one is worth a
longer warning because ARMA($1,1$) is so often used as a default without
checking whether it is actually identified by the data.

### Why near-cancellation matters

For an ARMA($1,1$) with our sign convention,

$$
\Phi(z) = 1 - \phi_1 z, \qquad \Theta(z) = 1 + \theta_1 z.
$$

If the two polynomials share a common root then the AR and MA operators
*cancel*. After cancellation the mean equation is just white noise around
$\mu$; the model has two parameters doing the job of zero and the likelihood
is flat along a ridge.  Even when the match is not exact, *near*-cancellation
makes the model practically unidentifiable: the data cannot tell whether the
observed persistence comes from the AR side, the MA side, or a mixture of both.

The fitted coefficients below illustrate exactly this.  The AR coefficient
is $\hat\phi_1 \approx -0.99$ and the MA coefficient is $\hat\theta_1 \approx
0.99$.  Because $\hat\phi_1 \approx -\hat\theta_1$, the two polynomials
$1 - \hat\phi_1 z$ and $1 + \hat\theta_1 z$ are almost identical, so their
roots are almost the same.  `arma_near_cancellation()` reports the distance
between the corresponding inverse roots.

## Demo

```{r}
library(tsgarch)
suppressMessages(library(data.table))
suppressMessages(library(xts))
data(nikkei)
nikkei <- xts(nikkei$value, as.Date(nikkei$index))
spec <- garch_modelspec(nikkei, arma = c(1,1), model = 'garch', constant = TRUE, 
                        init = 'unconditional', distribution = 'jsu')
mod <- estimate(spec)
as_flextable(summary(mod))
```

The estimated AR and MA coefficients and the inverse roots:

```{r}
arma_coefficients(mod)
arma_inverse_roots(mod)
arma_near_cancellation(arma_inverse_roots(mod), tol = 0.1)
```

The plot method now takes in an argument `type` which dispatches to either the GARCH (default and backwards compatible)
or ARMA plots.

```{r, fig.width=7,fig.height=6}
plot(mod)
```

```{r, fig.width=7,fig.height=6}
plot(mod, type = "arma", which = NULL, envelope = "parametric")
```

The first panel of the ARMA plot is the inverse-root diagram.  Because the
AR and MA inverse roots are extremely close, the plot joins them with a
dashed segment and the legend flags "Possible common factors".  This is the
visual cue that an ARMA($1,1$) mean equation may be over-parameterized for
this series.

