---
title: "Package Demo"
author: "Alexios Galanos"
date: "`r Sys.Date()`"
output: 
    rmarkdown::html_vignette:
        css: custom.css
        code_folding: show
vignette: >
  %\VignetteIndexEntry{Package Demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Estimation

We use the Deutsche Mark British Pound dataset for this demonstration throughout
the document.

```{r}
library(tsgarch)
suppressMessages(library(data.table))
suppressMessages(library(xts))
data(dmbp)
dmbp <- xts(dmbp, as.Date(1:nrow(dmbp), origin = '1970-01-01'))
spec <- garch_modelspec(dmbp[1:1500,1], model = 'fgarch', constant = TRUE, 
                        init = 'unconditional', distribution = 'jsu')
mod <- estimate(spec)
as_flextable(summary(mod))
```
Notice that the `as_flextable` method provides for a publication ready option
for printing out the model summary.

We next take a look at a summary plot of the estimated model:

```{r, fig.width=7,fig.height=6}
par(mar = c(2,2,2,2))
plot(mod)
```

Notice the news impact curve which is both shifted and rotated, a particularly
appealing feature of the Family GARCH model.


# Simulation and Prediction

The code below shows how to predict from an estimated model and how to simulate
using a specification with fixed parameters.

```{r, fig.width=7,fig.height=4}
delta <- coef(mod)["delta"]
new_spec <- spec
new_spec$parmatrix <- copy(mod$parmatrix)
sim <- simulate(new_spec, nsim = 500, h = 10000, seed = 100, burn = 1000)
mean_sim <- mean(rowMeans(sim$sigma^delta))^(2/delta)
pred <- predict(mod, h = 1000, nsim = 0)
par(mar = c(2,2,2,2))
plot(as.numeric(pred$sigma^2), type = "l", xlab = "horizon", 
     ylab = expression(sigma^2), ylim = c(0.25, 0.41), main = "Family GARCH - JSU Prediction")
abline(h = as.numeric(unconditional(mod)), col = 2)
abline(h = mean_sim, col = 3)
legend("bottomright", c("h-step prediction","unconditional variance","simulated unconditional variance"), col = c(1,2,3), lty = 1, bty = "n")
```


# Conclusion

There are other methods available such as one for profiling the parameter distribution (`tsprofile`),
a backtest method (`tsbacktest`) as well as other methods for extracting information from
the estimated object such as `vcov`, `pit` (probability integral transform), `confint` etc.

The package does not include any tests, as most of these have been moved to the 
[tstests](https://github.com/tsmodels/tstests/) package and there are ample packages 
which include testing methods.
